### 2018-05-08 article_springer/index.Rmd
set up for rmarkdown::render_site()

---
title: "Peaks Over Threshold for Bursty Time Series"
# titlerunning: "CTRE: POT for Bursty Time Series"
thanks: |
    Peter Straka was supported by the 
    Discovery Early Career Research Award DE160101147 on the Project 
    "Predicting Extremes when Events Occur in Bursts" by the Australian
    Research Council.
    Katharina Hees was supported by 
    the DAAD co-financed by the German Federal Ministry of Education and 
    Research (BMBF). 
authors:

- name: "Katharina Hees"
  address: University of Dortmund
  email: hees@statistik.uni-dortmund.de
- name: "Smarak Nayak"
  address: National Australia Bank
  email: smarak.nayak@nab.com.au
- name: "Peter Straka*"
  address: UNSW Sydney
  email: p.straka@unsw.edu.au 
authorrunning: K. Hees, S. Nayak & P. Straka
keywords:
- heavy tails
- renewal process
- extreme value theory
- peaks over threshold
abstract: |
  In many complex systems studied in statistical physics, 
  inter-arrival times between events such as solar flares, trades and 
  neuron voltages follow a heavy-tailed 
  distribution. The set of event times is fractal-like, being dense in some 
  time windows and empty in others, a phenomenon which has been dubbed 
  "bursty".

  This article generalizes the Peaks Over Threshold (POT) model to the setting 
  where inter-event times are heavy-tailed. For high thresholds and 
  infinite-mean waiting times, we show that the times between threshold 
  crossings are Mittag-Leffler distributed, and thus form a "fractional 
  Poisson Process" which generalizes the standard Poisson Process.
  We provide graphical means of estimating model parameters and assessing 
  model fit. Along the way, we apply our inference method to a real-world
  bursty time series, and 
  show how the memory of the Mittag-Leffler distribution affects 
  the predictive distribution for the time until the next extreme event. 
bibliography: CTRMstats.bib
output: 
  html_document: 
    toc: true
    toc_float: true
    theme: journal
  rticles::springer_article: 
    fig_caption: yes
params:
  tail: 0.8
  n: 10000
preamble: |
  \usepackage{amssymb}
---

```{r, child="article_springer.Rmd"}

```